---
title: "fastform"
output: html_document
---

```{r, echo = FALSE, include = FALSE}

  knitr::opts_chunk$set(warning = FALSE)
  knitr::opts_chunk$set(message = FALSE)
  knitr::opts_chunk$set(echo = FALSE)


  source("scripts/functions.R")

# Data ####

  source("scripts/data.R")

# Variables ####

  source("scripts/variables.R")

# Long datasets

  source("scripts/long.data.R")
  
```

### Fastform

## Bone healing

Study arm is not associated with the the status of the break at the end of the study.

```{r, include = FALSE}

  a <- with(data, table(outcome.final, arm,
                   dnn = c("Outcome", "Study Arm"))) 
 
```


```{r}
   CrossTable(a, digits = 2, prop.t = FALSE, prop.chisq = FALSE,
                 chisq = TRUE, fisher = TRUE) 
```

We can recode unknown outcomes as optimistic (everyone with an undetermined outcome evantually healed), pessimistic (they didn't heal). Either way, there is no difference between the groups. 

### Optimistic
```{r, include = FALSE}

  a <- with(data, table(outcome.op, arm,
                   dnn = c("Outcome", "Study Arm"))) 
 
```


```{r}
   CrossTable(a, digits = 2, prop.t = FALSE, prop.chisq = FALSE,
                 chisq = TRUE, fisher = TRUE) 
```

### Pessimistic
```{r, include = FALSE}

  a <- with(data, table(outcome.pes, arm,
                   dnn = c("Outcome", "Study Arm"))) 
 
```


```{r}
   CrossTable(a, digits = 2, prop.t = FALSE, prop.chisq = FALSE,
                 chisq = TRUE, fisher = TRUE) 
```

For other outcomes measured at 2 time points, there aren't enough observations at the end of the study to conduct any statistical tests. 

### Odour

No difference.

```{r, include = FALSE}

  a <- with(data, table(odour.early, arm,
                   dnn = c("Odour (Early)", "Study Arm"))) 
 
```


```{r}
   CrossTable(a, digits = 2, prop.t = FALSE, prop.chisq = FALSE,
                 chisq = TRUE, fisher = TRUE) 
```

Among those with previous cast experience. 

```{r, include = FALSE}

  a <- with(data[data$previous.cast.here.early == "yes", ],
            table(odour.early, arm,
            dnn = c("Odour (Early)", "Study Arm"))) 
 
```


```{r}
   CrossTable(a, digits = 2, prop.t = FALSE, prop.chisq = FALSE,
                 chisq = TRUE, fisher = TRUE) 
```


### Comfort

No difference.

```{r, include = FALSE}

  a <- with(data, table(comfort.early, arm,
                   dnn = c("Comfort (Early)", "Study Arm"))) 
 
```


```{r}
   CrossTable(a, digits = 2, prop.t = FALSE, prop.chisq = FALSE,
                 chisq = TRUE, fisher = TRUE) 
```

Among those with previous cast experience. 

```{r, include = FALSE}

  a <- with(data[data$previous.cast.here.early == "yes", ],
            table(comfort.early, arm,
            dnn = c("Comfort (Early)", "Study Arm"))) 
 
```


```{r}
   CrossTable(a, digits = 2, prop.t = FALSE, prop.chisq = FALSE,
                 chisq = TRUE, fisher = TRUE) 
```


### Waterproof rating

A clear difference favoring Fastform

```{r, include = FALSE}

  a <- with(data, table(waterproof.rate.early, arm,
                   dnn = c("Waterproof rating (Early)", "Study Arm"))) 
 
```


```{r}
   CrossTable(a, digits = 2, prop.t = FALSE, prop.chisq = FALSE,
                 chisq = TRUE, fisher = TRUE) 
```


## Quick Dash total scores

No differences.

```{r}

  select(quick.dash.long, qd, value.qd, arm, time.qd) %>%
   filter(qd == "qd.score") %>%
   mutate(time.qd = factor(time.qd, c("mid", "late"))) %>%
  ggplot(aes(x = value.qd, fill = arm, group = arm)) +
    geom_bar(color = "black", position = "dodge") +
    scale_fill_viridis("", discrete = TRUE) +
    theme_base() +
    facet_wrap(~arm + time.qd) +
    xlab("QuickDash score") +
    ylab("Participants")

```

### Wilcox rank tests

Early - No difference
```{r}
  wilcox.test(qd.score.mid ~ arm, data = data)

```

Late - No difference
```{r}
  wilcox.test(qd.score.late ~ arm, data = data)

```

### Quickdash items

```{r, include = FALSE}

  qd.comps <- select(data, id, arm, starts_with("qd.")) %>%
              select(everything(), -matches(".score.")) %>%
              select(everything(), -matches(".sport.")) %>%
              select(everything(), -matches(".work."))

  qd.comps[, c(3:24)] <- lapply(qd.comps[, c(3:24)],
                                function(x) factor(x,
                                                   c(1:5)))
```

```{r}
  filter(qd.comps, arm == "Fastform") %>%
    select(ends_with("mid")) %>%
    likert() %>%
    plot(type = "heat") +
    scale_fill_gradient2("Percent", low  = "white",
                                    mid  = viridis(1, begin = 1),
                                    high = viridis(1, begin = .5),
                         limits = c(0, 100)) +
    ggtitle("QuickDash items, Fastform, Week 7/8")

  filter(qd.comps, arm == "Resin") %>%
    select(ends_with("mid")) %>%
    likert() %>%
    plot(type = "heat") +
    scale_fill_gradient2("Percent", low  = "white",
                         mid  = viridis(1, begin = 1),
                         high = viridis(1, begin = .5),
                         limits = c(0, 100)) +
    ggtitle("QuickDash items, Resin, Week 7/8")

  filter(qd.comps, arm == "Fastform") %>%
    select(ends_with("late")) %>%
    likert() %>%
    plot(type = "heat") +
    scale_fill_gradient2("Percent", low  = "white",
                         mid  = viridis(1, begin = 1),
                         high = viridis(1, begin = .5),
                         limits = c(0, 100)) +
    ggtitle("QuickDash items, Fastform, Week 12/13")



  filter(qd.comps, arm == "Resin") %>%
    select(ends_with("late")) %>%
    likert() %>%
    plot(type = "heat") +
    scale_fill_gradient2("Percent", low  = "white",
                         mid  = viridis(1, begin = 1),
                         high = viridis(1, begin = .5),
                         limits = c(0, 100)) +
    ggtitle("QuickDash items, Resin, Week 12/13")
```


# EQ visual acuity scores

No differences
```{r}

 select(eq.long, eq, value.eq, arm, time.eq) %>%
    filter(eq == "eq.visual.acuity.score") %>%
    mutate(time.eq = factor(time.eq, c("two", "three"))) %>%
    mutate(value.eq = as.numeric(value.eq)) %>%
  ggplot(aes(x = value.eq, fill = arm, group = arm)) +
    geom_bar(color = "black", position = "dodge") +
    scale_fill_viridis("", discrete = TRUE) +
    theme_base() +
    facet_wrap(~arm + time.eq) +
    xlab("EQ visual aquity score") +
    ylab("Participants")

```

### Wilcox rank tests

Early - No difference

```{r}
  wilcox.test(eq.visual.acuity.score.two ~ arm, data = data)
```

Late - No difference

```{r}
  wilcox.test(eq.visual.acuity.score.three ~ arm, data = data)

```

### EQ items

```{r, include = FALSE}


 eq.comps <- select(data, id, arm, starts_with("eq.")) %>%
              select(everything(), -matches(".score."))
```

```{r}

 filter(eq.comps, arm == "Fastform") %>%
    select(ends_with("two")) %>%
    likert() %>%
    plot(type = "heat") +
    scale_fill_gradient2("Percent", low  = "white",
                         mid  = viridis(1, begin = 1),
                         high = viridis(1, begin = .5),
                         limits = c(0, 100)) +
    ggtitle("EQ items, Fastform, Week 2") +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))

  filter(eq.comps, arm == "Resin") %>%
    select(ends_with("two")) %>%
    likert() %>%
    plot(type = "heat") +
    scale_fill_gradient2("Percent", low  = "white",
                         mid  = viridis(1, begin = 1),
                         high = viridis(1, begin = .5),
                         limits = c(0, 100)) +
    ggtitle("EQ items, Resin, Week 2") +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))

  filter(eq.comps, arm == "Fastform") %>%
    select(ends_with("three")) %>%
    likert() %>%
    plot(type = "heat") +
    scale_fill_gradient2("Percent", low  = "white",
                         mid  = viridis(1, begin = 1),
                         high = viridis(1, begin = .5),
                         limits = c(0, 100)) +
    ggtitle("EQ items, Fastform, Week 3") +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))

  filter(eq.comps, arm == "Resin") %>%
    select(ends_with("three")) %>%
    likert() %>%
    plot(type = "heat") +
    scale_fill_gradient2("Percent", low  = "white",
                         mid  = viridis(1, begin = 1),
                         high = viridis(1, begin = .5),
                         limits = c(0, 100)) +
    ggtitle("EQ items, Resin, Week 3") +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

